---
title: "Gettign Started with the 'recent' Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This package implements the methods of [Fellows et. al. 2022](https://arxiv.org/abs/2204.00048) for the calculation of incidence using recency assays such as LAg-Avidity. Because treated individuals and elite controllers are known to ave high false positivity rates on recency assays, individuals with low viral load or a previous HIV+ diagnosis are screened out. A previous diagnosis can be indicated by either a self-reported previous positive test or through the detection of ARV antibodies.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Data

The package includes example data, which we will be using in this vignette. 

```{r setup}
library(recent)
data("assay_data")
head(assay_data[1:7])
```

1. `recent`: Logical. Did the subject test recent on the assay.
2. `hiv`: Logical. Is the subject HIV+.
3. `undiagnosed`: Logical. Has the subject never received a positive HIV test.
4. `elite_cntr`: Logical. Viral load test <1,000.
5. `tslt`: Numeric. Time since last HIV test in months.
6. `ever_hiv_test`: Logical. Has the subject never been tested for HIV.
7. `weights`: Numeric. Survey Weights.
8. `btwt001-btwt206`: Numeric. Jackknife replicate weights.

# Incidence Estimation

Incidence is calculated using the `rita_incidence` function. The `frr` and `assay_surv` are the reference FRR and reference probability of testing recent by day since seroconversion. Here we use values calculated for the LAg-Avidity assay. The default recency period (tau) is 2 years.

```{r}
inc <- rita_incidence(
    recent=assay_data$recent,
    undiagnosed=assay_data$undiagnosed,
    low_viral=assay_data$elite_cntr,
    hiv=assay_data$hiv,
    weights=assay_data$weights,
    tslt=assay_data$tslt,
    ever_hiv_test=assay_data$ever_hiv_test,
    frr = lag_avidity_frr()[1],
    assay_surv = lag_avidity_survival(2 * 365)
  )
 knitr::kable(inc)
```

The function outputs the following values:

1. `incidence`: The incidence.
2. `residual_frr`: The false recency rate accounting for the screening process.
3. `omega_rs`: The mean duration of recency up to tau accounting for the screening process.
4. `P(R|S)` : The proportion of screened in individual who test recent.
5. `P(S|H)` : The proportion of HIV+ individuals that are screened in.
6. `P(H)` : HIV prevalence.


By default the function estimates the time to diagnosis distribution using testing history information among undiagnosed HIV+ cases. If there are few of these, or their reliability is questionable, the HIV- population can be used instead by specifying ``.

```{r}
inc <- rita_incidence(
    recent=assay_data$recent,
    undiagnosed=assay_data$undiagnosed,
    low_viral=assay_data$elite_cntr,
    hiv=assay_data$hiv,
    weights=assay_data$weights,
    tslt=assay_data$tslt,
    ever_hiv_test=assay_data$ever_hiv_test,
    frr = lag_avidity_frr()[1],
    assay_surv = lag_avidity_survival(2 * 365),
    test_history_population = "negative"
  )
knitr::kable(inc)
```

# Bootstrap Confidence Intervals

Calculation of confidence intervals using survey replicate weights is supported. 

```{r}
rep_weights <-  dplyr::select(assay_data, dplyr::contains("btwt"))
ri <- rita_bootstrap(
    recent=assay_data$recent,
    undiagnosed=assay_data$undiagnosed,
    low_viral=assay_data$elite_cntr,
    hiv=assay_data$hiv,
    weights=assay_data$weights,
    tslt=assay_data$tslt,
    ever_hiv_test=assay_data$ever_hiv_test,
    rep_weights = rep_weights,
    rep_weight_type = "JK1"
  )
knitr::kable(ri)
```

Numerous jackknife and bootstrap weight types are supported. See `survey::svrrepdesign` for details.

